---
title: "Predictive Analytics"
author: "Rohin Mehra"
date: "5/13/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Preditictive Analytics


Loading libraries and reading file
```{r}
library(dplyr)
library(purrr)
library(tidyr)
library(data.table)
library(caret)
library(tidyverse)

dat<- read.csv("Life Expectancy Data.csv")
```


Checking for NA values and removing them

```{r pressure}
sum(is.na(dat))

#There are 2563 NA values.Removing them.

dat1<- na.omit(dat)
#Now no NA values in the dataset.
```

Lets consider Life expectancy to be our target variable.


Using Density Plot To Check If Response Variable Is Close To Normal
```{r plot}
library(e1071)  # for skewness function
par(mfrow=c(1, 2))  # divide graph area in 2 columns

plot(density(dat1$Life.expectancy), main="Density Plot: Life Expectancy", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(dat1$Life.expectancy), 2)))  # density plot for 'Life Expectancy'

polygon(density(dat1$Life.expectancy), col="red")

```


1. Removing the variables not to be used and partioning the data into Test ans Training set.

```{r}
dat2<- dat1[c(-1,-2,-3)]

set.seed(123)
training.samples <- dat2$Life.expectancy %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- dat2[training.samples, ]
test.data <- dat2[-training.samples, ]
```

80% of the data is partition into Training set and 20% in Test set.


2.Fitting the model

```{r Model}
model<- lm(Life.expectancy~., data = train.data)

summary(model)
```
Conclusion-
(a) From the model summary we can get that the R square value is 83%, whcihn is quite good. Also, p value is less than significant 0.05.
(b) The output shows the variables impacting the model. The variables having p value less than 0.05 are highly significant ones.
(c) Therefore removing the insignificant ones.


Model Update
```{r}
model_updated<- lm(Life.expectancy~Adult.Mortality+infant.deaths+percentage.expenditure+ under.five.deaths+
                     Total.expenditure+ Diphtheria+ HIV.AIDS+ Income.composition.of.resources+Schooling, data = train.data)

summary(model_updated)
```
Now all variables are highly significant.



3.Predicting the with test set-
```{r}
# Predict test data based on model
predict_reg <- model_updated %>% predict(test.data)

```

```{r}
linear_rmse<- data.frame(
  RMSE = caret::RMSE(predict_reg, test.data$Life.expectancy),
  Rsquare = caret::R2(predict_reg, test.data$Life.expectancy)
)

sigma(model_updated)/mean(test.data$Life.expectancy)
```
1. We can see that the RMSE value is low i.e 3.689 and R square is high i.e 81.8%, meaning model is good.
2. Also, the average prediction error rate is 5%.
